Optimal updating factor in Wang-Landau and metadynamics simulations

### 2016-04-16 doc/invt.tex
invt.tex: move a few sessions to Appendices

As shown above, with a fixed value of $q(T)$,
the optimization of the total error is equivalent to
that of the asymptotic part.
%
We can rewrite Eq. \eqref{eq:error_asym1} as an action
%
\begin{equation}
  \Err_A
  =
  \int_0^T
    {\mathcal L} \bigl[ q(t)\bigr]
    \, dt,
  \notag
  %\label{eq:error_asym_action}
\end{equation}
%
where the Lagrangian,
%
\begin{align}
  {\mathcal L} \bigl[ q(t) \bigr]
  &=
  \sum_{ k = 2 }^n
    \Gamma_k \, \dot u_k^2\bigl[ q(t) \bigr]
  \notag
  \\
  &=
  \sum_{ k = 2 }^n
    \Gamma_k \, \lambda_k^2 \, u_k^2 \bigl[ q(t) \bigr]
  \; \dot q^2( t ),
\notag
%\label{eq:error_asym_Lagrangian}
\end{align}
%
represents a free particle
with a position-dependent mass.
%
Since the Lagrangian
does not explicitly depend on time, $t$,
the Hamiltonian
%
\begin{equation}
  \mathcal H
  =
  \frac{ \partial \mathcal L }
       { \partial \dot q     }
  \, \dot q
  -
  \mathcal L
  =
  2 \, \mathcal L
  - \mathcal L
  =
  \mathcal L,
  %=
  %\mathrm{const.}
  \notag
  %\label{eq:error_asym_Hamiltonian}
\end{equation}
%
which coincides with the Lagrangian,
is conserved.
%
So we may set
%
\begin{equation}
  \mathcal L
  =
    \textstyle\sum_{ k = 2 }^n
      \Gamma_k \, \lambda_k^2
      \, u_k^2 \bigl[ q(t) \bigr]
  \;
  \dot q^2(t)
  =
  c^2
  ,
  \quad
  (c > 0).
  \label{eq:Lagrangian_const}
\end{equation}
%
Integrating the square root, we get
%
\begin{equation}
  \int_{ 0 }^{ q(t) }
    \sqrt{
      \textstyle\sum_{ k = 2 }^n
        \Gamma_k \, \lambda_k^2
        \, u_k^2( q' )
    }
    \;
    d q'
  =
  c \, t
  ,
  \label{eq:q_opt}
\end{equation}
%
where $c$ is to be determined by
the boundary condition at $t = T$ as
%
\begin{equation}
  c =
  \frac 1 T
  \int_{ 0 }^{ q(T) }
    \sqrt{
      \textstyle\sum_{ k = 2 }^n
        \Gamma_k \, \lambda_k^2
        \, u_k^2( q' )
    }
    \;
    d q'
  .
  \notag
  %\label{eq:q_opt}
\end{equation}
%
Thus, we have an implicit equation for $q(t)$,
and hence the optimal schedule,
$\alpha(t) = \dot q(t)$.
%
Since we only optimized the asymptotic error, $E_A$,
in th above,
the parameter $q(T)$ can be further adjusted
to minimize the total error, $E = E_R + E_A$.

@@

\subsubsection{\label{sec:measure_gamma}
Measurement of the integrals
of the autocorrelation functions}

To use Eq. \eqref{eq:q_opt} in a real simulation,
we need to know both $\lambda_k$'s and $\Gamma_k$'s.
%
While the updating scheme alone
determines the eigenmodes and $\lambda_k$'s,
the values of $\Gamma_k$'s depend also on the system
and the underlying sampling process (MC or MD).
%
Below is a technique of measuring $\Gamma_k$'s.

\subsubsection{\label{sec:program}
Program}

We can now give a complete program of computing
the optimal schedule for a given updating scheme.

The dependence of variables are shown in Fig. \ref{fig:vardep}.

\begin{figure}
\begin{tikzpicture}
  \node[draw, text width={width("Updating ")}]
    (w) at (0, 3) {Updating \\matrix, $\mathbf w$};

\end{tikzpicture}
\caption{
  \label{fig:vardep}
  Dependence of variables in computing
  the optimal schedule.
}
\end{figure}

@@

Note that Eq. \eqref{eq:wband_eigenvalue_refl}
differs from Eq. \eqref{eq:wband_eigenvalue_pbc}
by a factor of $2$
in the argument of the cosine function.
%
We may unify the two as
%
\begin{equation}
  \lambda_k
  =
  \mu_0
  +
  2
  \sum_{ l = 1 }^b
    \mu_l \,
    \cos \frac{ (k - 1) \, l \, g \, \pi }
              {            n             }
  ,
  \label{eq:wband_eigenvalue}
\end{equation}
%
where the degeneracy $g$ is $2$
for the periodic case,
or $1$ otherwise.
%
In both cases,
the eigenvalues, $\lambda_k$'s,
is the cosine transform of
the updating kernel.
%
For the updating scheme to be stable,
the cosine transform of the kernel
must be nonnegative.
%
\note{If this is not so,
we can modify the kernel
using the technique described
in Sec. \ref{sec:stabilize_wband}.}

@@

For $n \gg 1$,
we define
$x = l \, g \, \Delta x$,
and
$\mu_l = \mu(x) \, \Delta x$,
where
$\Delta x = \pi/n$
is the bin size.
%
Then,
by assuming the largest $b \approx n/g$,
Eq. \eqref{eq:wband_eigenvalue}
can be approximated as an integral:
%
\begin{equation}
  \lambda_k
  =
  2 \int_0^\pi
    \mu(x) \, \cos \bigl[ (k-1) \, x \bigr] \, dx,
\notag
%\label{eq:lambda_int}
\end{equation}
%
with the normalization
%
\begin{equation}
  1 = 2 \int_0^\pi \mu(x) \, dx.
\notag
%\label{eq:mx_normalization}
\end{equation}

@@

\subsubsection{\label{sec:stabilize_wband}
Inversion and Stabilization}

%A practical problem of using homogeneous updating scheme
%is the following.
%
A homogeneous updating scheme
specified by an arbitrary kernel
is not necessarily stable
with all nonnegative eigenvalues.
%
%
\note{For example,
  %the nearest-neighbor updating scheme
  %is unstable if Eq. \eqref{eq:nn_stable} is violated.
  %
  %Even
  the Gaussian updating scheme can be unstable
  if the updating kernel is truncated.
}%
%
Below we show how to minimally modify
the updating kernel
to stabilize it.
%
\note{The relative code is \texttt{trimwindow()} in \texttt{invt.h}.
}

The key is the observation of Eq. \eqref{eq:wband_eigenvalue}
that the eigenvalues are related to the updating kernel
by a cosine transform,
and thus the relation can be readily inverted.
%
In the periodic case,
the inversion formula is
%
\begin{equation}
  \mu_l
  =
  \frac 1 n
  \sum_{ k = 1 }^n
  \lambda_k
  \cos \frac{ (k - 1) \, l \, 2 \, \pi }
            {            n             }
  .
\label{eq:mu_from_lambda_pbc}
\end{equation}

@@

Now given an updating kernel,
we can compute the eigenvalues from
Eq. \eqref{eq:wband_eigenvalue}.
%
By replacing negative eigenvalues with zeros and
using Eq. \eqref{eq:mu_from_lambda_pbc}
or \eqref{eq:mu_from_lambda_refl},
we get a new updating kernel,
which is stable by construction.
%
However, the new kernel is widened
as $\mu_l$ may be nonzero for $l > b$.
%
If we truncate the new kernel at $\mu_b$,
the stability problem remains,
although the negative eigenvalues
often become smaller in magnitude.
%
Thus, to stabilize the updating kernel
without increasing the width,
we need to iterate the above process many times,
resulting the following algorithm.

@@

In the above, we have focused on
the asymptotic error.
%
For a finite-length simulation,
we should include the residual part
and minimize the total error.
%
For the bandpass updating schemes,
this leads to the optimal value
of the free parameter, $t_0$.
%
%Assuming the system is initially equilibrated
%under a constant updating magnitude,
%$\alpha_0$,
Assuming Eq. \eqref{eq:error_res1} with
$$
q(T)
=
\int_0^T \frac{ dt } { t + t_0 }
=
\ln \frac{ T + t_0 } { t_0 }.
$$
%
we have, from Eqs. \eqref{eq:error_split}
and \eqref{eq:error_asym_sinc},
%
\begin{equation}
  E
  =
  E_R + E_A
  =
  \frac { \frac 1 2 \alpha_0 \, t_0^2 + T }
        {          ( T + t_0 )^2          }
  \sum_{ k = 2 }^K \Gamma_k
  .
\notag
%\label{eq:error_sinc0}
\end{equation}
%
The minimum is reached at
%
\begin{equation}
  t_0
  =
  \frac{     2    }
       { \alpha_0 }
  ,
\label{eq:t0_sinc}
\end{equation}
%
and at this point
%
\begin{equation}
  E
  =
  \frac{   1     }
       { T + t_0 }
  \sum_{ k = 2 }^K
    \Gamma_k
  .
\label{eq:error_sinc}
\end{equation}
%
At the beginning of the schedule,
the updating magnitude is
$\alpha(0^+) = \alpha_0 / 2$,
half of the previous equilibrium value.
%
This factor $1/2$
happens to coincide with the
reduction factor of updating magnitude
at stage transitions
in the original WL algorithm\cite{
wang2001, wang2001pre}.

@@

\subsubsection{\label{sec:equilerr}
Comparison to the error of the equilibrium FDS method}

@@

Each adaptive FDS method defines an updating scheme,
or mathematically an updating matrix,
whose elements $w_{ij}$ specifies the
relative magnitude of updating the bias potential
at bin $i$ upon a visit to bin $j$.
%
Our method computes the optimal schedule from
the eigenvalues of the updating matrix,
and the integrals of the correlation functions
the corresponding eigenmodes.
%
For the single-bin (WL) scheme,
the optimal schedule from our method
is given by inverse-time formula,
Eq. \eqref{eq:alpha_invt1}.
%
The parameter, $t_0$, can be adjusted
according to Eq. \eqref{eq:t0_sinc},
to take into account the initial error.
%such that the initial updating magnitude
%is half of the previous equilibrium value.
%
For a general multiple-bin scheme,
including the Gaussian one used in metadynamics,
the optimal schedule has no closed form,
but is given by an equation of motion
of a free particle with a position-dependent mass, Eq. \eqref{eq:q_opt}.
%
In this case,
the optimal schedule and error
can be sensitive to the simulation length.

@@

@@

For the one-step sampling process,
we find from Eq. \eqref{eq:Gamma_onestep}
that $\Gamma(p) = \cot^2 p$
is peaked around $p \approx 0$,
%
Then
$$
  \Err_A
  \propto
  \frac{   2 \, n }
       { \pi \, T }
  \frac{             1             }
       {  \lambda \, (2 - \lambda) }
  .
$$
\note{
$$
\begin{aligned}
  \Err_A
  &
  \propto
  \frac{   2 \, n }
       { \pi \, T }
  \left.
  \frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
       { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
  \right|_{ p \approx 0 }
  \notag \\
  &
  \approx
  \frac{   2 \, n }
       { \pi \, T }
  \frac{             1             }
       {  \lambda \, (2 - \lambda) }
  .
\end{aligned}
$$
}%
%
The optimal $\lambda$ is therefore
%
\begin{equation}
\lambda \approx 1.
\label{eq:lambda_nn_onestep}
\end{equation}
%
\note{We can be a bit more precise here.
%
From Eq. \eqref{eq:Gamma_onestep}, we get
%
$$
\Gamma(p) = \cot^2 p \approx \cos^2 p / (\sin^2 p + \delta^2),
$$
%
where we have introduced
$\delta \propto \sin[ \pi / (2 \, n) ] \sim 1/n$
as a small parameter to avoid the divergence
around the origin.
%
$$
\begin{aligned}
\Err_A
&=
\frac{   2 \, n }
     { \pi \, t }
\int_0^{\pi/2}
    \frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
         { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
    \frac{ \cos^2 p }
         { \sin^2 p + \delta^2 }
\, dp
\notag \\
&
\stackrel{    \delta \ll 1     }
         { =\joinrel=\joinrel= }
\frac{   2 \, n }
     { \pi \, t }
\frac{             1             }
     {  \lambda \, (2 - \lambda) }
\frac{    1   }
     { \delta }
.
\end{aligned}
$$
The integral is evaluated as follows.
%
We again change variable $p \to \frac{ \pi } { 2 } - p$,
and
$$
\begin{aligned}
\Err_A
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\int_0^{2 \, \pi}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
     {       2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
\cdot
\frac{ \sin^2 p            }
     { \cos^2 p + \delta^2 }
\, dp
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[1 -      \mu_1 \, \left(z + \frac{1}{z}\right)^2 \right]^2 }
     {       2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
\cdot
\frac{ 4 - \left( z + \frac 1 z \right)^2 }
     { \left( z + \frac 1 z \right)^2 + 4 \, \delta^2 }
\, \frac{dz}{z}
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[z^2 -      \mu_1 \, (z^2 + 1)^2 \right]^2   }
     {  (2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2 }
\cdot
\frac{ 4 \, z^2 - ( z^2 + 1 )^2 }
     { ( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2 }
\, \frac{ dz }{ z^3 }
\\
&=
\frac{ n } { \lambda \, t }
\sum_l \operatorname{Res}_l
,
\end{aligned}
$$
where the contour is the unit circle
in the complex plane of $z$,
and the integral is reduced to a sum of residues
around the poles within.

@@

>>>>

As shown above, with a fixed value of $q(T)$,
the optimization of the total error is equivalent to
that of the asymptotic part.
%
We can rewrite Eq. \eqref{eq:error_asym1} as an action
%
\begin{equation}
  \Err_A
  =
  \int_0^T
    {\mathcal L} \bigl[ q(t)\bigr]
    \, dt,
  \notag
  %\label{eq:error_asym_action}
\end{equation}
%
where the Lagrangian,
%
\begin{align}
  {\mathcal L} \bigl[ q(t) \bigr]
  &=
  \sum_{ k = 2 }^n
    \Gamma_k \, \dot u_k^2\bigl[ q(t) \bigr]
  \notag
  \\
  &=
  \sum_{ k = 2 }^n
    \Gamma_k \, \lambda_k^2 \, u_k^2 \bigl[ q(t) \bigr]
  \; \dot q^2( t ),
\notag
%\label{eq:error_asym_Lagrangian}
\end{align}
%
represents a free particle
with a position-dependent mass.
%
Since the Lagrangian
does not explicitly depend on time, $t$,
the Hamiltonian
%
\begin{equation}
  \mathcal H
  =
  \frac{ \partial \mathcal L }
       { \partial \dot q     }
  \, \dot q
  -
  \mathcal L
  =
  2 \, \mathcal L
  - \mathcal L
  =
  \mathcal L,
  %=
  %\mathrm{const.}
  \notag
  %\label{eq:error_asym_Hamiltonian}
\end{equation}
%
which coincides with the Lagrangian,
is conserved.
%
So we may set
%
\begin{equation}
  \mathcal L
  =
    \textstyle\sum_{ k = 2 }^n
      \Gamma_k \, \lambda_k^2
      \, u_k^2 \bigl[ q(t) \bigr]
  \;
  \dot q^2(t)
  =
  c^2
  ,
  \quad
  (c > 0).
  \label{eq:Lagrangian_const}
\end{equation}
%
Integrating the square root, we get
%
\begin{equation}
  \int_{ 0 }^{ q(t) }
    \sqrt{
      \textstyle\sum_{ k = 2 }^n
        \Gamma_k \, \lambda_k^2
        \, u_k^2( q' )
    }
    \;
    d q'
  =
  c \, t
  ,
  \label{eq:q_opt}
\end{equation}
%
where $c$ is to be determined by
the boundary condition at $t = T$ as
%
\begin{equation}
  c =
  \frac 1 T
  \int_{ 0 }^{ q(T) }
    \sqrt{
      \textstyle\sum_{ k = 2 }^n
        \Gamma_k \, \lambda_k^2
        \, u_k^2( q' )
    }
    \;
    d q'
  .
  \notag
  %\label{eq:q_opt}
\end{equation}
%
Thus, we have an implicit equation for $q(t)$,
and hence the optimal schedule,
$\alpha(t) = \dot q(t)$.
%
Since we only optimized the asymptotic error, $E_A$,
in th above,
the parameter $q(T)$ can be further adjusted
to minimize the total error, $E = E_R + E_A$.
%
A complete program of computing the optimal schedule
is listed in Appendix \ref{sec:program}.

@@

@@

Note that Eq. \eqref{eq:wband_eigenvalue_refl}
differs from Eq. \eqref{eq:wband_eigenvalue_pbc}
by a factor of $2$
in the argument of the cosine function.
%
We may unify the two as
%
\begin{equation}
  \lambda_k
  =
  \mu_0
  +
  2
  \sum_{ l = 1 }^b
    \mu_l \,
    \cos \frac{ (k - 1) \, l \, g \, \pi }
              {            n             }
  ,
  \label{eq:wband_eigenvalue}
\end{equation}
%
where the degeneracy $g$ is $2$
for the periodic case,
or $1$ otherwise.
%
In both cases,
the eigenvalues, $\lambda_k$'s,
is the cosine transform of
the updating kernel.
%
For the updating scheme to be stable,
the cosine transform of the kernel
must be nonnegative.
%
If this is not so,
we can modify the kernel
using the technique described
in Appendix \ref{sec:stabilize_wband}.

@@

For $n \gg 1$,
we define
$x = l \, g \, \Delta x$,
and
$\mu_l = \mu(x) \, \Delta x$,
where
$\Delta x = \pi/n$
is the bin size.
%
Then,
by assuming the largest $b \approx n/g$,
Eq. \eqref{eq:wband_eigenvalue}
can be approximated as an integral:
%
\begin{equation}
  \lambda_k
  =
  2 \int_0^\pi
    \mu(x) \, \cos \bigl[ (k-1) \, x \bigr] \, dx,
\notag
%\label{eq:lambda_int}
\end{equation}
%
with the normalization
%
$
  1 = 2 \int_0^\pi \mu(x) \, dx.
$

@@

\subsubsection{\label{sec:invert_wband}
Inversion}

Since the eigenvalues are related to the updating kernel
by a cosine transform,
Eq. \eqref{eq:wband_eigenvalue},
the relation can be readily inverted.
%
In the periodic case,
we have
%the inversion formula is
%
\begin{equation}
  \mu_l
  =
  \frac 1 n
  \sum_{ k = 1 }^n
  \lambda_k
  \cos \frac{ (k - 1) \, l \, 2 \, \pi }
            {            n             }
  .
\label{eq:mu_from_lambda_pbc}
\end{equation}

@@

@@

In the above, we have focused on
the asymptotic error.
%
For a finite-length simulation,
we should include the residual part
and minimize the total error.
%
For the bandpass updating schemes,
this leads to the optimal value
of the free parameter, $t_0$.
%
%Assuming the system is initially equilibrated
%under a constant updating magnitude,
%$\alpha_0$,
Assuming Eq. \eqref{eq:error_res1} with
$$
q(T)
=
\int_0^T \frac{ dt } { t + t_0 }
=
\ln \frac{ T + t_0 } { t_0 }.
$$
%
we have, from Eqs. \eqref{eq:error_split}
and \eqref{eq:error_asym_sinc},
%
\begin{equation}
  E
  =
  E_R + E_A
  =
  \frac { \frac 1 2 \alpha_0 \, t_0^2 + T }
        {          ( T + t_0 )^2          }
  \sum_{ k = 2 }^K \Gamma_k
  .
\notag
%\label{eq:error_sinc0}
\end{equation}
%
The minimum is reached at
%
\begin{equation}
  t_0
  =
  \frac{     2    }
       { \alpha_0 }
  ,
\label{eq:t0_sinc}
\end{equation}
%
and at this point
%
\begin{equation}
  E
  =
  \frac{   1     }
       { T + t_0 }
  \sum_{ k = 2 }^K
    \Gamma_k
  .
\label{eq:error_sinc}
\end{equation}
%
In Appendix \ref{sec:equilerr},
we show that the above error tends
to the error of the ideal equilibrium FDS simulation
in the asymptotic limit.

Note also that at the beginning of the schedule,
the updating magnitude is
$\alpha(0^+) = \alpha_0 / 2$,
half of the previous equilibrium value.
%
This factor $1/2$
happens to coincide with the
reduction factor of updating magnitude
at stage transitions
in the original WL algorithm\cite{
wang2001, wang2001pre}.

@@

@@

Each adaptive FDS method defines an updating scheme,
or mathematically an updating matrix,
whose elements $w_{ij}$ specifies the
relative magnitude of updating the bias potential
at bin $i$ upon a visit to bin $j$.
%
Our method computes the optimal schedule from
the eigenvalues of the updating matrix,
and the integrals of the correlation functions
the corresponding eigenmodes.
%
For the single-bin (WL) scheme,
the optimal schedule from our method
is given by inverse-time formula,
Eq. \eqref{eq:alpha_invt1}.
%
The parameter, $t_0$, can be adjusted
according to Eq. \eqref{eq:t0_sinc},
to accommodate the initial error.
%such that the initial updating magnitude
%is half of the previous equilibrium value.
%
For a general multiple-bin scheme,
including the Gaussian one used in metadynamics,
the optimal schedule has no closed form,
but is given by an equation of motion
of a free particle with a position-dependent mass, Eq. \eqref{eq:q_opt}.
%
In this case,
the optimal schedule and error
can be sensitive to the simulation length.

@@

\section{\label{sec:program}
Program of computing the optimal schedule and error
}

Here we give a program of computing
the optimal schedule and the error
of a given adaptive FDS simulation.
%
Basically, we extract
both $\lambda_i$'s and $\Gamma_i$'s
from the input,
and then feed them to Eq. \eqref{eq:q_opt}
for the schedule.
%
The dependence of variables are shown in Fig. \ref{fig:vardep}.

\begin{figure}[h]
\begin{tikzpicture}
  \node[draw, text width={width("Updating ")}]
    (w) at (0, 3) {Updating \\matrix, $\mathbf w$};

\end{tikzpicture}
\caption{
  \label{fig:vardep}
  Dependence of variables in computing
  the optimal schedule and the error.
}
\end{figure}

%\subsection{\label{sec:measure_gamma}
%Measurement of the integrals
%of the autocorrelation functions}

%To use Eq. \eqref{eq:q_opt} in a real simulation,
%we need to know both $\lambda_k$'s and $\Gamma_k$'s.
%
While the updating scheme alone
determines the eigenmodes and $\lambda_k$'s,
the values of $\Gamma_k$'s depend also on the system
and the underlying sampling process (MC or MD).
%
Below is a technique of measuring $\Gamma_k$'s.

%\subsection{\label{sec:program_list}
%Program
%}

The complete program is the following.

@@

For the one-step sampling process,
we find from Eq. \eqref{eq:Gamma_onestep}
that $\Gamma(p) = \cot^2 p$
is peaked around $p \approx 0$,
%
Then
$$
  \Err_A
  \propto
  \frac{   2 \, n }
       { \pi \, T }
  \frac{             1             }
       {  \lambda \, (2 - \lambda) }
  .
$$
\note{
$$
\begin{aligned}
  \Err_A
  &
  \propto
  \frac{   2 \, n }
       { \pi \, T }
  \left.
  \frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
       { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
  \right|_{ p \approx 0 }
  %\notag \\
  %&
  \approx
  \frac{   2 \, n }
       { \pi \, T }
  \frac{             1             }
       {  \lambda \, (2 - \lambda) }
  .
\end{aligned}
$$
}%
%
The optimal $\lambda$ is, therefore,
%
\begin{equation}
\lambda \approx 1.
\label{eq:lambda_nn_onestep}
\end{equation}
%
\note{We can be a bit more precise here.
%
From Eq. \eqref{eq:Gamma_onestep}, we get
%
$$
\Gamma(p) = \cot^2 p \approx \cos^2 p / (\sin^2 p + \delta^2),
$$
%
where we have introduced
$\delta \propto \sin[ \pi / (2 \, n) ] \sim 1/n$
as a small parameter to avoid the divergence
around the origin.
%
$$
\begin{aligned}
\Err_A
&=
\frac{   2 \, n }
     { \pi \, t }
\int_0^{\pi/2}
    \frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
         { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
    \frac{ \cos^2 p }
         { \sin^2 p + \delta^2 }
\, dp
\notag \\
&
\stackrel{    \delta \ll 1     }
         { =\joinrel=\joinrel= }
\frac{   2 \, n }
     { \pi \, t }
\frac{             1             }
     {  \lambda \, (2 - \lambda) }
\frac{    1   }
     { \delta }
.
\end{aligned}
$$
The integral is evaluated as follows.
%
We again change variable $p \to \frac{ \pi } { 2 } - p$,
and
$$
\begin{aligned}
\Err_A
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\int_0^{2 \, \pi}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
     {       2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
\cdot
\frac{ \sin^2 p            }
     { \cos^2 p + \delta^2 }
\, dp
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[1 -      \mu_1 \, \left(z + \frac{1}{z}\right)^2 \right]^2 }
     {       2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
\cdot
\frac{ 4 - \left( z + \frac 1 z \right)^2 }
     { \left( z + \frac 1 z \right)^2 + 4 \, \delta^2 }
\, \frac{dz}{z}
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[z^2 -      \mu_1 \, (z^2 + 1)^2 \right]^2   }
     {  (2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2 }
\cdot
\frac{ 4 \, z^2 - ( z^2 + 1 )^2 }
     { ( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2 }
\, \frac{ dz }{ z^3 }
\\
&=
\frac{ n } { \lambda \, t }
\sum_l \operatorname{Res}_l
,
\end{aligned}
$$
where the contour is the unit circle
in the complex plane of $z$,
and the integral is reduced to a sum of residues
around the poles within.

@@

\section{\label{sec:stabilize_wband}
Stabilization of homogeneous updating schemes}

%A practical problem of using homogeneous updating scheme
%is the following.
%
A homogeneous updating scheme
specified by an arbitrary kernel
is not necessarily stable
with all nonnegative eigenvalues.
%
%
\note{For example,
  %the nearest-neighbor updating scheme
  %is unstable if Eq. \eqref{eq:nn_stable} is violated.
  %
  %Even
  the Gaussian updating scheme can be unstable
  if the updating kernel is truncated.
}%
%
Here we show how to minimally modify
the updating kernel
to stabilize it.
%
\note{The relative code is \texttt{trimwindow()} in \texttt{invt.h}.
}

Given an updating kernel,
we can compute the eigenvalues from
Eq. \eqref{eq:wband_eigenvalue}.
%
By replacing negative eigenvalues with zeros and
using Eq. \eqref{eq:mu_from_lambda_pbc}
or \eqref{eq:mu_from_lambda_refl},
we get a new updating kernel,
which is stable by construction.
%
However, the new kernel is widened
as $\mu_l$ may be nonzero for $l > b$.
%
If we truncate the new kernel at $\mu_b$,
the stability problem remains,
although the negative eigenvalues
often become smaller in magnitude.
%
Thus, to stabilize the updating kernel
without increasing the width,
we need to iterate the above process many times,
resulting the following algorithm.

\section{\label{sec:equilerr}
Comparison to the error of the equilibrium FDS simulation}
